### 2015-02-27 paper/paper_conclusion.tex

% !TEX root = paper.tex
% !TEX encoding = UTF-8 Unicode

\section{Conclusion}

Our results indicate that, at least for this language pair and this data set, post-editing quality is consistently higher, by a statistically significant amount, when bilingual post-editors are presented with alignment data.
%
%Our results indicate that this finding holds across participants and across segments of varying MT quality. 
%
This result holds across participants, and is strongest when MT quality is poor.