### 2015-02-27 paper/paper_results.tex

% !TEX root = paper.tex
% !TEX encoding = UTF-8 Unicode

\input{figure2}

\section{Results}
\label{sec:results}

%\subsection{Rating of translations}

An experienced English-Russian translator and grader rated all post-edited segments as well as the MT segments.%
%
\footnote{We include these adequacy judgements, along with the other data and code produced or used in this work, in the attached dataset and software supplements.}
%in both human-readable tabular form and machine-readable CSV form in the attached dataset supplement.}
%
The rater was presented with blocks of 10 lines arranged in a vertical list:
%
the first line was a source segment, 
%
the second was its reference translation, and 
%
below these were eight translations for the segment 
%
(the MT, one monolingual post-edited translation from \citet{2014_WMT_Schwartz_etal}, and six bilingual post-edited translations) 
%
listed in random order. 
%
The rater was asked to classify all the segment translations on an adequacy scale (see Table \Vref{judge_guidelines}) ranging from 2 (translation makes no sense) to 12 (translation is superior to the reference translation).

\todo[inline]{Add Spanish paragraph here}

%The post-edited English segments

The mean adequacy score when bilingual participants were presented with alignments was 8.35.
%
When alignments were omitted from the post-editing tool, the mean adequacy score was 7.85.  
%
A Wilcoxon signed-rank test \citep{1945_Wilcoxon} showed that when participants were presented with alignments the ratings of their translations were significantly higher than when participants post-edited without access to alignments (N = 6, Z = -2.207, p = 0.027).

\todo[inline]{Add Spanish paragraph here}

This finding indicates that, at least for this language pair and data set, bilingual post-editors produce higher quality translations when presented with bilingual alignment links between source words and machine-translated target words.

In addition to rating each post-edited translation, our bilingual human judge rated the adequacy of the unedited MT segments.
%
We thereby obtained direct human judgements of MT quality.%
%
\footnote{Because we care about the adequacy of post-edited translations, we consider actual human judgements to be preferable to automated metrics such as BLEU \citep{2002_ACL_Papineni_etal}, which at best serve as a flawed proxy for human judgements.}
%
We partition the data according to the MT quality as measured by our human judge in terms of adequacy.
%
For each partition, we then calculate the mean adequacy score of post-edited segments, once for the case where post-editors were presented with alignment data, and once for the case where alignments were not presented.
%
In addition, we perform the same average calculation for the monolingual post-edited segments of Docs A and B from \citet{2014_WMT_Schwartz_etal}.
%
The results are shown in Figure \Vref{fig:mean_adequacy_score}.

%\clearpage

\input{figure3}

\input{figure4}

We observe that when MT quality was poor (adequacy categories 2--6), bilingual post-editors consistently produced higher quality post-edited translations when alignments were shown.
%
When MT quality was high (adequacy categories 8--10), little or no impact to adequacy quality was observed when alignments were shown. 
%
By subtracting the adequacy score of each machine translated segment, we obtain the adequacy gain obtained by post-editing.
%
The above effect can be seen clearly in Figure \Vref{fig:mean_adequacy_gain}, where we show mean gain in adequacy over unedited MT;
%
post-editing makes the most difference when MT quality is poor, and in those cases the presence of alignment links improves post-editing as measured by adequacy.

In Figure \ref{fig:mean_adequacy_score_per_posteditor}, we examine the effect that alignment link visualization has on each bilingual post-editor.
%
We observe that post-editing quality varies widely between post-editors (with PE2 and PE3 performing best).
%
For all six bilingual post-editors, we observe higher mean adequacy scores when alignment links were presented than when they were omitted from the post-editing tool.
%
We also note that when alignment links were absent, one bilingual post-editor (PE5) performed worse than the monolingual post-editor (PE0) from \citet{2014_WMT_Schwartz_etal}.
%
When compared to the unedited machine translations, post-editing resulted in improved mean adequacy for all post-editors, both bilingual and monolingual.

Finally, in Figure \ref{fig:percentage_segments} we present histograms graphing the percentage of segments in each adequacy category.
%
We observe that the majority of unedited machine translations are of relatively poor quality, falling mostly into categories 4 and 6. %, and a few scoring as 12.
%
%The monolingual post-editor from \citet{2014_WMT_Schwartz_etal} was able to skew the quality histogram somewhat to the right, with most segments scored as 6 or 10, and a few scored as better than the reference (12).
The monolingual post-editor improved MT quality somewhat, with most segments scoring as 6 or 10.
%
The histogram for bilingual post-editors using no alignments skewed further to the right, peaking at 8.
%
The histogram for bilingual post-editors with access to alignments skews farthest to the right, peaking at 10.