### 2015-02-27 paper/paper_title.tex

% !TEX root = paper.tex
% !TEX encoding = UTF-8 Unicode

\title{The Effect of Word Alignments in Bilingual and Monolingual Post-Editing}

\author{%
        {}\\
	    {}\\
	    {}\\
	    {}\\
	    {\tt}
}

\date{}