### 2015-07-10 paper/paper_title.tex

% !TEX root = paper.tex
% !TEX encoding = UTF-8 Unicode

%\title{Effects of Word Alignment Visualization on Post-Editing Quality \& Speed}
%
%\author{{}\\
%	    {}\\
%	    {}\\
%	    {}\\
%	    {\tt}
%}
%
%\date{}

\title{\bf Effects of Word Alignment Visualization on Post-Editing Quality \& Speed}

\author{\name{\bf First Author} \hfill  \addr{author1@abc.university.country}\\ 
        \name{\bf Second Author} \hfill \addr{author2@abc.university.country}\\ 
        \addr{Department of Science, My University, 
        MyTown, Zip, Country}
\AND
       \name{\bf Third Author} \hfill \addr{author3@abc.university.country}\\
        \addr{Department of Science, My University, 
        MyTown, Zip, Country}
}