### 2015-09-07 paper/paper_title.tex

% !TEX root = paper.tex
% !TEX encoding = UTF-8 Unicode

%\title{Effects of Word Alignment Visualization on Post-Editing Quality \& Speed}
%
%\author{{}\\
%	    {}\\
%	    {}\\
%	    {}\\
%	    {\tt}
%}
%
%\date{}

\title{\bf Effects of Word Alignment Visualization on Post-Editing Quality \& Speed}

\author{\name{\bf Lane Schwartz} \hfill  \addr{lanes@illinois.edu}\\ 
        \addr{Dept. of Linguistics, University of Illinois at Urbana-Champaign, Urbana, IL 61801, USA}
\AND
        \name{\bf Isabel Lacruz} \hfill \addr{ilacruz@kent.edu}\\ 
       \name{\bf Tatyana Bystrova} \hfill \addr{tbystro2@kent.edu}\\
        \addr{Dept. of Modern \& Classical Language Studies, Kent State University, Kent, OH 44242, USA}
}